Tutorial 3b

### 2025-07-14 Tutorial_1a.Rmd
update Tutorials and ms.

Load our custom functions. Make sure you specify the same levels and labels for factor condition in the function **plot_eha** (line 164-166 in custom_functions.R) as above (lines 119-122).

>>>>

Load our custom functions. Make sure you specify the same levels and labels for factor condition in the function **plot_eha** (line 164-166 in custom_functions.R) as above (lines 124-126).

### 2025-07-14 Tutorial_2a.Rmd
update Tutorials and ms.

M0d took about 77 minutes.

@@

Model_M0c took about 19 minutes to run.

@@

Model M1i took about 124 minutes to fit.

@@

Second, when the shape of the hazard function is rather smooth, as it is for behavioral RT data, one can fit a more parsimonious model by treating TIME as a continuous variable, and use a polynomial specification of the effect of TIME. Thus, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach for prime-target congruency (variable "condition") and treat TIME as a continuous variable (variable "period_9").

@@

The specification "0 + Intercept + ..." removes the default intercept in brm() and adds an explicit Intercept for which we can set the prior ourselves. The variable period_9_sq is a squared version of period_9. Note that duplicate terms in the model formula (e.g., condition) are ignored.
Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model_M1d took about 184 minutes to run.

@@

  ggplot(aes(x=bin, y=diff, fill=contrast)) +
     stat_lineribbon(alpha = 0.5, 
                  point_interval = "mean_qi",
                  .width=c(.8,.95)) +
    geom_hline(yintercept = 0, 
               color = "red", 
               lty = 3) +
    scale_y_continuous(limits=c(-0.5,0.3)) +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    theme(axis.text.x = element_text(angle=90)) +
    labs(y = "difference in predicted hazard", x = "timebin") +
    ggtitle("Grand average marginal effect") +
    facet_wrap(~contrast)
```

@@

The shape of the prior predictive distribution for 1 bin looks as expected (compare with Supplementary Figure 2 (row F) in the Supplemental Material): hazard values below .5 are more likely than hazard values above .5.

>>>>

M0d took about 77 minutes using the same MacBook Pro.

@@

Model_M0c took about 19 minutes to run using the same MacBook Pro.

@@

Model M1i took about 124 minutes to fit using the same MacBook Pro.

@@

Second, when the shape of the hazard function is rather smooth, one can fit a more parsimonious model by treating TIME as a continuous variable, and use a polynomial specification of the effect of TIME. Thus, if we want to make assumptions about (1) how hazard changes over TIME in the reference condition (blank prime), and (2) how the effect of congruent and incongruent primes change over TIME (relax the proportionality assumption), then we can switch to a dummy coding approach for prime-target congruency (variable "condition") and treat TIME as a continuous variable (variable "period_9").

@@

The specification "0 + Intercept + ..." removes the default intercept in brm() and adds an explicit Intercept for which we can set the prior ourselves. The variable period_9_sq is a squared version of period_9. Note that duplicate terms in the model formula (e.g., condition) are ignored.
Because TIME is centered on bin 9, the Intercept represents the estimated cloglog-hazard in bin 9 for the blank prime condition. Model_M1d took about 184 minutes to run using the same MacBook Pro.

@@

  ggplot(aes(x=bin, y=diff, fill=contrast)) +
     stat_lineribbon(alpha = 0.5, 
                  point_interval = "mean_qi",
                  .width=c(.8,.95)) +
    geom_hline(yintercept = 0, 
               color = "red", 
               lty = 3) +
    scale_y_continuous(limits=c(-0.5,0.3)) +
    scale_x_discrete(labels = str_c("(",(c(6:15)-1)*40,",",c(6:15)*40,"]"), breaks = 6:15) +
    scale_fill_brewer(palette = "Dark2") +
    scale_color_brewer(palette = "Dark2") +
    theme(axis.text.x = element_text(angle=90)) +
    labs(y = "difference in predicted hazard", x = "timebin") +
    facet_wrap(~contrast)
```

@@

The shape of the prior predictive distribution for 1 bin looks as expected (compare with Supplementary Figure 3 (row F) in section F of the Supplemental Material): hazard values below .5 are more likely than hazard values above .5.

### 2025-07-14 Tutorial_3a.Rmd
update Tutorials and ms.

Model_M1i_f took about 3 hours.

@@

# plot
p1 <- ggplot(data = est, aes(x = position)) +
 geom_point(aes( y = brm[,"Estimate"],  
                 color = "c1", 
                 shape = "s1"),
                 size = 2.4) +
 geom_point(aes( y = glmer$estimate, 
                 color = "c2", 
                 shape = "s2"),
                 size = 2.4, 
                 position = position_nudge(x = .4))  +
 geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
                    color = "c1"), 
                show.legend = F,
                linewidth = 1.2) +
 geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, 
                    color = "c2"), 
                show.legend = F, 
                linewidth = 1.2,
                position = position_nudge(x = .4)) +
# geom_hline(yintercept=0,colour = "grey", linetype = 2) +
 scale_x_continuous(breaks = 6:15, labels = 6:15*40) +
 scale_y_continuous(limits=c(-9,3), breaks = c(-9,-6,-3,0,3)) +

 labs(x = "bin endpoint", y = "cloglog-hazard") +
 scale_color_manual(name = "Model", 
                    breaks = c("c1", "c2"), 
                    values = cols,
                    labels = c("brm", "glmer")) +
 scale_shape_manual(name = "Model", 
                    breaks = c("s1", "s2"),
                    values = shapes,
                    labels = c("brm", "glmer")) +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle=90)) +
  facet_wrap(~prime)
p1
```

>>>>

Model_M1i_f took about 3 hours on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).

@@

# plot
p1 <- ggplot(data = est, aes(x = position)) +
 geom_point(aes( y = brm[,"Estimate"],  
                 color = "c1", 
                 shape = "s1"),
                 size = 2.4) +
 geom_point(aes( y = glmer$estimate, 
                 color = "c2", 
                 shape = "s2"),
                 size = 2.4, 
                 position = position_nudge(x = .4))  +
 geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
                    color = "c1"), 
                show.legend = F,
                linewidth = 1.2) +
 geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, 
                    color = "c2"), 
                show.legend = F, 
                linewidth = 1.2,
                position = position_nudge(x = .4)) +
 scale_x_continuous(breaks = 6:15, labels = 6:15*40) +
 scale_y_continuous(limits=c(-9,3), breaks = c(-9,-6,-3,0,3)) +
 labs(x = "time bin (upper bound in ms)", y = "cloglog-hazard") +
 scale_color_manual(name = "Model", 
                    breaks = c("c1", "c2"), 
                    values = cols,
                    labels = c("brm", "glmer")) +
 scale_shape_manual(name = "Model", 
                    breaks = c("s1", "s2"),
                    values = shapes,
                    labels = c("brm", "glmer")) +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle=90)) +
  facet_wrap(~prime)
p1
```

### 2025-07-14 Tutorial_3b.Rmd
update Tutorials and ms.

Model_M1i_ca_f took about 90 minutes to fit.

@@

# plot
p1 <- ggplot(data = est, aes(x = position)) +
 geom_point(aes( y = brm[,"Estimate"],  
                 color = "c1", 
                 shape = "s1"),
                 size = 2.4) +
 geom_point(aes( y = glmer$estimate, 
                 color = "c2", 
                 shape = "s2"),
                 size = 2.4, 
                 position = position_nudge(x = .4))  +
 geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
                    color = "c1"), 
                linewidth = 1.2,
                show.legend = F) +
 geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, 
                    color = "c2"), 
                show.legend = F,
                linewidth = 1.2,
                position = position_nudge(x = .4)) +
 #geom_hline(yintercept=0,colour = "grey", linetype = 2) +
 scale_x_continuous(breaks = 6:15, labels = 6:15*40) +
 scale_y_continuous(limits=c(-25,25), breaks = c(-25,-20,-15,-10,-5,0,5,10,15,20,25)) +

 labs(x = "bin endpoint", y = "logit-ca") +
 scale_color_manual(name = "Model", 
                    breaks = c("c1", "c2"), 
                    values = cols,
                    labels = c("brm", "glmer")) +
 scale_shape_manual(name = "Model", 
                    breaks = c("s1", "s2"),
                    values = shapes,
                    labels = c("brm", "glmer")) +
 theme(panel.grid.minor = element_blank(),
       axis.text.x = element_text(angle=90)) +
facet_wrap(~prime)
p1
```

>>>>

Model_M1i_ca_f took about 90 minutes to fit on a MacBook Pro (Sonoma 14.6.1 OS, 18GB Memory, M3 Pro Chip).

@@

# plot
p1 <- ggplot(data = est, aes(x = position)) +
 geom_point(aes( y = brm[,"Estimate"],  
                 color = "c1", 
                 shape = "s1"),
                 size = 2.4) +
 geom_point(aes( y = glmer$estimate, 
                 color = "c2", 
                 shape = "s2"),
                 size = 2.4, 
                 position = position_nudge(x = .4))  +
 geom_linerange(aes(ymin=brm[,"Q2.5"], ymax=brm[,"Q97.5"], 
                    color = "c1"), 
                linewidth = 1.2,
                show.legend = F) +
 geom_linerange(aes(ymin=glmer$conf.low, ymax=glmer$conf.high, 
                    color = "c2"), 
                show.legend = F,
                linewidth = 1.2,
                position = position_nudge(x = .4)) +
 scale_x_continuous(breaks = 6:15, labels = 6:15*40) +
 scale_y_continuous(limits=c(-25,25), breaks = c(-25,-20,-15,-10,-5,0,5,10,15,20,25)) +
 labs(x = "time bin (upper bound in ms)", y = "logit-ca") +
 scale_color_manual(name = "Model", 
                    breaks = c("c1", "c2"), 
                    values = cols,
                    labels = c("brm", "glmer")) +
 scale_shape_manual(name = "Model", 
                    breaks = c("s1", "s2"),
                    values = shapes,
                    labels = c("brm", "glmer")) +
 theme(panel.grid.minor = element_blank(),
       axis.text.x = element_text(angle=90)) +
facet_wrap(~prime)
p1
```
